group

Manage groups. On most platforms this can only create groups. Group membership must be managed on individual users.

On some platforms such as OS X, group membership is managed as an attribute of the group, not the user record. Providers must have the feature ‘manages_members’ to manage the ‘members’ property of a group record.
Features

    manages_aix_lam: The provider can manage AIX Loadable Authentication Module (LAM) system.
    manages_members: For directories where membership is an attribute of groups not users.
    system_groups: The provider allows you to create system groups with lower GIDs.

Provider 	manages_aix_lam 	manages_members 	system_groups
aix 	X 	X 	 
directoryservice 	  	X 	 
groupadd 	  	  	X
ldap 	  	  	 
pw 	  	  	 
Parameters
allowdupe

Whether to allow duplicate GIDs. This option does not work on FreeBSD (contract to the pw man page). Valid values are true, false.
attribute_membership

Whether specified attribute value pairs should be treated as the only attributes of the user or whether they should merely be treated as the minimum list. Valid values are inclusive, minimum.
attributes

Specify group AIX attributes in an array of keyvalue pairs Requires features manages_aix_lam.
auth_membership

whether the provider is authoritative for group membership.
ensure

Create or remove the group. Valid values are present, absent.
gid

The group ID. Must be specified numerically. If not specified, a number will be picked, which can result in ID differences across systems and thus is not recommended. The GID is picked according to local system standards.
ia_load_module

The name of the I&A module to use to manage this user Requires features manages_aix_lam.
members

The members of the group. For directory services where group membership is stored in the group objects, not the users. Requires features manages_members.
name

The group name. While naming limitations vary by system, it is advisable to keep the name to the degenerate limitations, which is a maximum of 8 characters beginning with a letter.
provider

The specific backend for provider to use. You will seldom need to specify this — Puppet will usually discover the appropriate provider for your platform. Available providers are:

    aix: Group management for AIX! Users are managed with mkgroup, rmgroup, lsgroup, chgroup Required binaries: /usr/sbin/lsgroup, /usr/sbin/rmgroup, /usr/bin/chgroup, /usr/bin/mkgroup. Default for operatingsystem == aix. Supported features: manages_aix_lam, manages_members.

    directoryservice: Group management using DirectoryService on OS X.

    Required binaries: /usr/bin/dscl. Default for operatingsystem == darwin. Supported features: manages_members.

    groupadd: Group management via groupadd and its ilk.

    The default for most platforms

    Required binaries: groupdel, groupmod, groupadd. Supported features: system_groups.

    ldap: Group management via ldap.

    This provider requires that you have valid values for all of the ldap-related settings, including ldapbase. You will also almost definitely need settings for ldapuser and ldappassword, so that your clients can write to ldap.

    Note that this provider will automatically generate a GID for you if you do not specify one, but it is a potentially expensive operation, as it iterates across all existing groups to pick the appropriate next one.

    pw: Group management via pw.

    Only works on FreeBSD.

    Required binaries: /usr/sbin/pw. Default for operatingsystem == freebsd.

system

Whether the group is a system group with lower GID. Valid values are true, false.
