user

Manage users. This type is mostly built to manage system users, so it is lacking some features useful for managing normal users.

This resource type uses the prescribed native tools for creating groups and generally uses POSIX APIs for retrieving information about them. It does not directly modify /etc/passwd or anything.

Autorequires: If Puppet is managing the user’s primary group (as provided in the gid attribute), the user resource will autorequire that group. If Puppet is managing any role accounts corresponding to the user’s roles, the user resource will autorequire those role accounts.
Features

    allows_duplicates: The provider supports duplicate users with the same UID.
    manages_aix_lam: The provider can manage AIX Loadable Authentication Module (LAM) system.
    manages_expiry: The provider can manage the expiry date for a user.
    manages_homedir: The provider can create and remove home directories.
    manages_password_age: The provider can set age requirements and restrictions for passwords.
    manages_passwords: The provider can modify user passwords, by accepting a password hash.
    manages_solaris_rbac: The provider can manage roles and normal users
    system_users: The provider allows you to create system users with lower UIDs.

Provider 	allows_duplicates 	manages_aix_lam 	manages_expiry 	manages_homedir 	manages_password_age 	manages_passwords 	manages_solaris_rbac 	system_users
aix 	  	X 	X 	X 	X 	X 	  	 
directoryservice 	  	  	  	  	  	X 	  	 
hpuxuseradd 	X 	  	  	X 	  	  	  	 
ldap 	  	  	  	  	  	X 	  	 
pw 	X 	  	  	X 	  	  	  	 
user_role_add 	X 	  	  	X 	X 	X 	X 	 
useradd 	X 	  	X 	X 	X 	X 	  	X
Parameters
allowdupe

Whether to allow duplicate UIDs. Valid values are true, false.
attribute_membership

Whether specified attribute value pairs should be treated as the only attributes of the user or whether they should merely be treated as the minimum list. Valid values are inclusive, minimum.
attributes

Specify user AIX attributes in an array of keyvalue pairs Requires features manages_aix_lam.
auth_membership

Whether specified auths should be treated as the only auths of which the user is a member or whether they should merely be treated as the minimum membership list. Valid values are inclusive, minimum.
auths

The auths the user has. Multiple auths should be specified as an array. Requires features manages_solaris_rbac.
comment

A description of the user. Generally is a user’s full name.
ensure

The basic state that the object should be in. Valid values are present, absent, role.
expiry

The expiry date for this user. Must be provided in a zero padded YYYY-MM-DD format - e.g 2010-02-19. Requires features manages_expiry.
gid

The user’s primary group. Can be specified numerically or by name.
groups

The groups of which the user is a member. The primary group should not be listed. Multiple groups should be specified as an array.
home

The home directory of the user. The directory must be created separately and is not currently checked for existence.
ia_load_module

The name of the I&A module to use to manage this user Requires features manages_aix_lam.
key_membership

Whether specified key value pairs should be treated as the only attributes of the user or whether they should merely be treated as the minimum list. Valid values are inclusive, minimum.
keys

Specify user attributes in an array of keyvalue pairs Requires features manages_solaris_rbac.
managehome

Whether to manage the home directory when managing the user. Valid values are true, false.
membership

Whether specified groups should be treated as the only groups of which the user is a member or whether they should merely be treated as the minimum membership list. Valid values are inclusive, minimum.
name

User name. While limitations are determined for each operating system, it is generally a good idea to keep to the degenerate 8 characters, beginning with a letter.
password

The user’s password, in whatever encrypted format the local machine requires. Be sure to enclose any value that includes a dollar sign ($) in single quotes (‘). Requires features manages_passwords.
password_max_age

The maximum amount of time in days a password may be used before it must be changed Requires features manages_password_age.
password_min_age

The minimum amount of time in days a password must be used before it may be changed Requires features manages_password_age.
profile_membership

Whether specified roles should be treated as the only roles of which the user is a member or whether they should merely be treated as the minimum membership list. Valid values are inclusive, minimum.
profiles

The profiles the user has. Multiple profiles should be specified as an array. Requires features manages_solaris_rbac.
project

The name of the project associated with a user Requires features manages_solaris_rbac.
provider

The specific backend for provider to use. You will seldom need to specify this — Puppet will usually discover the appropriate provider for your platform. Available providers are:

    aix: User management for AIX! Users are managed with mkuser, rmuser, chuser, lsuser Required binaries: /usr/sbin/lsgroup, /bin/chpasswd, /usr/sbin/lsuser, /usr/sbin/rmuser, /usr/bin/chuser, /usr/bin/mkuser. Default for operatingsystem == aix. Supported features: manages_aix_lam, manages_expiry, manages_homedir, manages_password_age, manages_passwords.
    directoryservice: User management using DirectoryService on OS X. Required binaries: /usr/bin/dscl. Default for operatingsystem == darwin. Supported features: manages_passwords.
    hpuxuseradd: User management for hp-ux! Undocumented switch to special usermod because HP-UX regular usermod is TOO STUPID to change stuff while the user is logged in. Required binaries: /usr/sam/lbin/userdel.sam, /usr/sam/lbin/usermod.sam, /usr/sbin/useradd. Default for operatingsystem == hp-ux. Supported features: allows_duplicates, manages_homedir.

    ldap: User management via ldap. This provider requires that you have valid values for all of the ldap-related settings, including ldapbase. You will also almost definitely need settings for ldapuser and ldappassword, so that your clients can write to ldap.

    Note that this provider will automatically generate a UID for you if you do not specify one, but it is a potentially expensive operation, as it iterates across all existing users to pick the appropriate next one. Supported features: manages_passwords.
    pw: User management via pw on FreeBSD. Required binaries: pw. Default for operatingsystem == freebsd. Supported features: allows_duplicates, manages_homedir.
    user_role_add: User management inherits useradd and adds logic to manage roles on Solaris using roleadd. Required binaries: roleadd, roledel, userdel, passwd, rolemod, usermod, useradd. Default for operatingsystem == solaris. Supported features: allows_duplicates, manages_homedir, manages_password_age, manages_passwords, manages_solaris_rbac.
    useradd: User management via useradd and its ilk. Note that you will need to install the Shadow Password Ruby library often known as ruby-libshadow to manage user passwords. Required binaries: userdel, chage, usermod, useradd. Supported features: allows_duplicates, manages_expiry, manages_homedir, manages_password_age, manages_passwords, system_users.

role_membership

Whether specified roles should be treated as the only roles of which the user is a member or whether they should merely be treated as the minimum membership list. Valid values are inclusive, minimum.
roles

The roles the user has. Multiple roles should be specified as an array. Requires features manages_solaris_rbac.
shell

The user’s login shell. The shell must exist and be executable.
system

Whether the user is a system user with lower UID. Valid values are true, false.
uid

The user ID. Must be specified numerically. For new users being created, if no user ID is specified then one will be chosen automatically, which will likely result in the same user having different IDs on different systems, which is not recommended. This is especially noteworthy if you use Puppet to manage the same user on both Darwin and other platforms, since Puppet does the ID generation for you on Darwin, but the tools do so on other platforms.
