vcsrepo

A local version control repository
Features

    bare_repositories: The provider differentiates between bare repositories and those with working copies
    filesystem_types: The provider supports different filesystem types
    gzip_compression: The provider supports explicit GZip compression levels
    reference_tracking: The provider supports tracking revision references that can change over time (eg, some VCS tags and branch names)

Provider 	bare_repositories 	filesystem_types 	gzip_compression 	reference_tracking
bzr 	  	  	  	X
cvs 	  	  	X 	X
git 	X 	  	  	X
hg 	  	  	  	X
svn 	  	X 	  	X
Parameters
compression

Compression level Requires features gzip_compression.
ensure

Valid values are present, bare, absent, latest.
fstype

Filesystem type Requires features filesystem_types.
path

    namevar

Absolute path to repository
provider

The specific backend for provider to use. You will seldom need to specify this — Puppet will usually discover the appropriate provider for your platform. Available providers are:

    bzr: Supports Bazaar repositories Required binaries: bzr. Default for bzr == exists. Supported features: reference_tracking.
    cvs: Supports CVS repositories/workspaces Required binaries: cvs. Default for cvs == exists. Supported features: gzip_compression, reference_tracking.
    git: Supports Git repositories Required binaries: git. Default for git == exists. Supported features: bare_repositories, reference_tracking.
    hg: Supports Mercurial repositories Required binaries: hg. Default for hg == exists. Supported features: reference_tracking.
    svn: Supports Subversion repositories Required binaries: svn, svnadmin. Default for svn == exists. Supported features: filesystem_types, reference_tracking.

revision

The revision of the repository Values can match /^\S+$/.
source

The source URI for the repository
